brados: declarative, programmable object storage

### 2016-05-24 paper.tex
paper update
Signed-off-by: Noah Watkins <noahwatkins@gmail.com>

%%% why is it hard / why do naive approaches fail
In addition to the growth in the quantity of operations in use throughout Ceph
installations, Figure~\ref{fig:objclass-dev} also depicts the amount of
low-level C++ written to implement these operations. Unfortunately, this code
is written assuming a performance profile defined by the combination of the
hardware and software versions available at the time of development.  While
the bulk of these interfaces are created by core Ceph developers with a
complete view of the performance model, this may be changing as the
development community has been receptive to outside contributions with the
recent inclusion by CERN developers of an extension for performing limited
numeric operations on object data~\cite{cls_numops}. And while Ceph has not
yet reached the point of directly exposing these features to
non-administrative users, the recent inclusion of a mechanism for dynamically
defining extensions using Lua~\cite{cls_lua} suggests that aspects of this
feature may soon appear. What is needed is support for creating storage
interfaces using a method that allows transparent optimization as the system,
application, and supporting environment evolve.

% why hasn't it been solved before / why have previous approaches failed
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the later has
been specifically concerned with the creation of application-specific
interfaces and demonstrating short-term benefits, the standardization seeks to
develop a last set of interfaces. In neither case have we seen efforts focused
on extensible interfaces with portability and longevity a primary concern.

%This trend is also not limited to Ceph. AWS Lambda, Redis Lua, Redis Modules,
%KV-Drives, and Rhea, are all recent examples of storage systems embracing the
%power of including application-specific semantics.

We propose the introduction of a declarative language for building object
based interfaces that allows a storage system to meet the needs of an
application throughout the development process without requiring rewrites.

@@

\section{Physical Design}

@@

As Ceph continues to develop into a target for building distributed services
and application, it is important to not restrict use to developers seasonsed
in distributed programming and those with knowledge of intracacies of the Ceph
performance model.  To this end we present Brados, a declarative language with
relational semantics for defining the behavior of RADOS object interfaces.

The Brados model corresponds to a subset of the Bloom language, and models the
storage system state uniformly as a collection of
relations~\cite{alvaro:cidr11}. The composition of a collection of existing
interfaces is then expressed as a collection of high-level \emph{queries} that
describe how a stream of requests (API calls) are filtered, transformed and
combined with existing state to define streams of outputs (API call returns as
well as updates to system state).  Separating the relational semantics of such
compositions from details of their physical implementations introduces a
number of degrees of freedom, similar to the ``data independence’’ offered by
database query languages.  The choice of access methods (for example, whether
to use a bytestream interface or a key/value store), storage decide
classes(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
1:1 or 1:N mapping) and execution plans (e.g.  operator ordering) can be
postponed to execution time.  The optimal choices for these physical design
details are likely to change much more often than the logical specification,
freeing the interface designer from the need to rewrite systems and device and
interface characteristics change.

>>>>

%%% why is it hard / why do naive approaches fail
In addition to the growth in the quantity of operations in use throughout Ceph
installations, Figure~\ref{fig:objclass-dev} also depicts the amount of
low-level C++ written to implement these operations. Unfortunately, this code
is written assuming a performance profile defined by the combination of the
hardware and software versions available at the time of development.  While
the bulk of these interfaces are created by core Ceph developers with a
complete view of the performance model, this may be changing as the
development community has been receptive to outside contributions with the
recent inclusion by CERN developers of an extension for performing limited
numeric operations on object data~\cite{cls_numops}. And while Ceph has not
yet reached the point of directly exposing these features to
non-administrative users, the under review inclusion of a mechanism for dynamically
defining extensions using Lua~\cite{cls_lua} suggests that aspects of this
feature may soon appear. What is needed is support for creating storage
interfaces using a method that allows transparent optimization as the system,
application, and supporting environment evolve.

% why hasn't it been solved before / why have previous approaches failed
Unfortunately the size of the design space for even simple interfaces can
be very large, and it can be difficult to choose a design that future proofs an
implementation against future hardware and software upgrades.
Previous work related to storage interface design has largely been in the
context of standardization efforts and active storage. While the former is
primarily concerned with fixed interfaces, application-specific interfaces
built using active storage techniques have not to the best of our knoweldge
addressed portability concerne that arise from defining interfaces
imperatively. We address this gap by using a declarative language we call
\emph{brados} to define application-specific storage interfaces in
object-based storage systems, decoupling interface definition from
implementation and allowing interfaces to adapt to system changes without
modification.

To demonstrate the use of the \emph{brados} language we will show how a
high-performance shared-log based on the CORFU protocol can be built, as well
as existing real-world interfaces used in the Ceph storage system. First we
discuss the Ceph storage system which our prototype is built upon and the 
motivating system, CORFU. We then present the challenges that programmers face
when navigating the design space during the process of building an
application-storage interface. The \emph{brados} language is then described
using the CORFU system and an existing Ceph-specific interface as examples.
We conclude by discussing the optimization opportunities that we can realize
using analysis of interfaces. I hope you enjoy the show.

@@

\section{Physical Design Space}

@@

As Ceph continues to develop into a target for building distributed services
and application, it is important to not restrict use to developers seasonsed
in distributed programming and those with knowledge of intracacies of the Ceph
performance model. To this end we present Brados, a declarative language with
relational semantics for defining the behavior of RADOS object interfaces.
Brados uses a simple declarative data model and can reduce the learning curve
for new users, and allows existing developers to increase productivity by
writing less code that is more portable.

The Brados language corresponds to a subset of the Bloom language, and models the
storage system state uniformly as a collection of
relations~\cite{alvaro:cidr11}. The composition of a collection of existing
interfaces is then expressed as a collection of high-level \emph{queries} that
describe how a stream of requests (API calls) are filtered, transformed and
combined with existing state to define streams of outputs (API call returns as
well as updates to system state).  Separating the relational semantics of such
compositions from details of their physical implementations introduces a
number of degrees of freedom, similar to the ``data independence’’ offered by
database query languages.  The choice of access methods (for example, whether
to use a bytestream interface or a key/value store), storage decide
classes(e.g., whether to use HDDs or SDDs), physical layout details (e.g. a
1:1 or 1:N mapping) and execution plans (e.g.  operator ordering) can be
postponed to execution time.  The optimal choices for these physical design
details are likely to change much more often than the logical specification,
freeing the interface designer from the need to rewrite systems and device and
interface characteristics change.
